A Multi-telescope Campaign on \frb: Implications for the FRB Population

### 2017-04-27 repeater-vladata.tex
waiting on jim and kevin additions

We used a Markov chain Monte Carlo (MCMC) method to model the dynamic radio spectra (time versus frequency). We used the Goodman and Weare affine invariant sampler \citep{goodman2010ensemble} as implemented in \emph{emcee} \citep{2013PASP..125..306F}. The dynamic spectra were modeled prior to dispersion correction. The spectral structure was modeled as a broad Gaussian envelope (parameters of amplitude, center, width), while the time evolution was modeled by a $\nu^2$\ dispersion law (parameters of starting integration, DM). Finally, the pulses were assumed to have an intrinsically square temporal width (parameter W). This 6-parameter model defined the intrinsic signal that was then distributed over a fixed time-frequency observing grid. The inter-channel amplitude modulation was ignored in the model.

@@

The parameters for the best model of the dynamic spectrum are given in Table \ref{tab:spec} and the resulting Gaussian model overlaid on Figure \ref{fig:spec}. The typical burst spectrum has a FWHM of 500~MHz. All but two of the best-fit Gaussians are centered inside the 3~GHz band and some appear contained by the 1~GHz wide band. This is consistent with previous detections of \frb\ by Arecibo, which showed quasi-broadband structure \citep{2016arXiv160308880S} and large variation in the implied spectral index \citep{2014ApJ...790..101S}.

% ** quantify how much of model shape is actually inside the band **

% ** confusion as to whether inter-channel modulation is signal or noise ** jim also to do noise analysis **
% ** could quantify deviations from Gaussian model relative to noise. An SNR relative to envelope.

% ** spectra plot modifications: **
% y-axis scale not shown! 
% add mjd to each plot
% Need to make sure impression differs of noise and inter-channel modulation

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57649.png}
 \end{minipage}
\caption{Spectra and best-fit model for nine bursts seen by the VLA from 2.5 to 3.5~GHz in the peak, dedispersed 5~ms integration. Starting at the top left (moving right and then down), they correspond to bursts on MJDs 57623, 57633.68, 57633.70, 57638, 57643, 57645, 57646, 57648, and 57649. The solid line is a best-fit Gaussian model found through modeling.
\label{fig:spec}}
\end{center}
\end{figure*}

@@

% ** fix plot to show "W" instead of "dint"

\begin{figure*}[htb]
\begin{center}
\includegraphics[width=2\columnwidth]{corner57646}
\caption{A scatterplot matrix that shows the correlation of every pair of parameters in the MCMC run for burst 57646. The equivalent plot for the other bursts show normally-distributed samples, but this burst shows significant structure in the samples. The parameters ``Amp'', ``Center'', and ``$\sigma_{ch}$'' refer to the Gaussian spectral shape, while ``DM'', ``int'', and ``W'' refer to the temporal shape. The parameters are given in units of channels (4 MHz) and integrations (5 ms).
\label{fig:corner}}
\end{center}
\end{figure*}

@@

Figure \ref{fig:ed} shows the \frb\ burst energy cumulative distribution as seen by the VLA and calculated from prior observations by Arecibo and the GBT \citep{2016Natur.531..202S,2016arXiv160308880S}. The latter two energy distributions are scaled from the fluence by assuming that all burst energy is included by the observation. This likely underestimates the burst energy for some bursts by a factor of a few, but should not affect the slope of the distribution.

@@

The burst-specific DM structure could be intrinsic to the emission mechanism or induced during propagation. Radio waves can be modified in a variety of ways (e.g., scintillation, scattering) and the duration of the emission in the source frame is not known \citep{2016arXiv160505890C}. \citet{CORDES} describe a model where plasma lensing near the source of an FRB can magnify radio emission by orders of magnitude. That model argues that the simultaneous appearance of a burst from 1.2 to 3.5~GHz requires a lens with the focal frequency be $\gtrsim 3.5$~GHz. This gives a constraint on the combination 
$$\frac{\rm DM_0}{a_{\rm AU}^2} \frac{d_{sl}d_{lo}/d_{so}}{1\, \rm kpc} \gtrsim (3.5 / 39.1)^2$$
where DM$_0$, $a$, and $d_{sl, lo, so}$\ are the electron column density of the lens, size of the lens, and lensing distances, respectively. A lens placed at $d_{sl}\approx1$~kpc is consistent with typical properties of the interstellar medium. Lenses that are influenced by the source would be much nearer to the source $d_{\sl} \ll 1$~kpc, requiring large values of $\rm{DM_0}/a_{\rm AU}^2$.

@@

% ** discuss beaming **

Our refined analysis shows that \frb\ has even higher isotropic energies than previously reported for \frb, reaching energies $E_{\rm{max}}\approx10^{40}$\ erg. \frb\ has isotropic energies orders of magnitude higher than the strongest Crab giant pulses ($\sim10^{35}$~erg) and comparable brightness temperatures \citep[$T_{b,\rm{Crab}}\sim10^{41}$~K versus $T_{b, \rm{FRB 121102}}\sim10^{38}$~K;][]{2003Natur.422..141H,2014PhRvD..89j3009K}. 
%While this emission is coherent and strongly beamed \citep{2016Natur.531..202S, WEIRD}, the energy scale is still larger than any known Galactic radio transient and requires either a new process or dramatic scaling of known emission processes \citep{2016MNRAS.462..941L, 2016MNRAS.457..232C}.

@@

%While the \frb\ burst energy distribution is consistent with the overall FRB population, this simulation ignores many observational biases (scattering, observing coverage) and the effects of galaxy evolution, which is likely significant over the range of redshifts to which \frb\ is detectable \citep[e.g.,][]{2010ApJ...709..644I}. The burst energies are estimated from observations from 2.5 to 3.5~GHz, which may be a biased selection of all bursts or may not include all burst emission. If the simulation is revealing a connection between \frb\ and the broader FRB population, it does not explain the physical mechanism that drives it. The burst energy distribution of the population could be dominated by intrinsic or extrinsic effects \citep{2015MNRAS.451.3278M,2016Natur.531..202S,CORDES}. More FRB detections will improve the quality of this inference as sample variance drops, while more burst detetions from \frb\ will improve the measurement of the intrinsic energy distribution.

@@

The chance of detecting a burst from \frb\ with the VLA changes substantially on day--month timescales. A major outstanding question is whether this time-variable rate is driven by an intrinsic process \citep{2016ApJ...826..226K} or extrinsic (propagation) effects in the host galaxy or intergalactic medium \citep{CORDES}. However, the temporal distribution of VLA detections alone shows that the \frb\ burst rate is significantly correlated on short timescales. That is, detecting a burst predicts that you are more likely to find another burst soon thereafter and not detecting a burst predicts you will likely not detect another burst soon thereafter.

@@

Using an assumed horizon redshift of 1, we estimate a rate of $R_{\rm{FRB}} \approx 5\times10^{-5} N_r^{-1} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$. This calculation ignores time dilation, which underestimates the burst rate by a factor of 2 at the redshift horizon. The narrow burst spectral structure seen for \frb\ also suggests that the FRB rate is understated by a factor of a few by ignoring bursts that fall outside the observing band. Finally, we note that this estimate assumes that Galactic latitude and propagation effects do not significantly affect the detectability of FRBs, although scintillation and scattering likely play significant roles \citep{2015MNRAS.451.3278M, CORDES}.

>>>>

We used a Markov chain Monte Carlo (MCMC) method to model the dynamic radio spectra (time versus frequency). We used the Goodman and Weare affine invariant sampler \citep{goodman2010ensemble} as implemented in \emph{emcee} \citep{2013PASP..125..306F}. The dynamic spectra were modeled prior to dispersion correction. The spectral structure was modeled as a broad Gaussian envelope (parameters of amplitude, center, $\sigma$), while the time evolution was modeled by a $\nu^2$\ dispersion law (parameters of starting integration, DM). Finally, the pulses were assumed to have an intrinsically square temporal width (parameter W). This 6-parameter model defined the intrinsic signal that was then distributed over a fixed time-frequency observing grid. The inter-channel amplitude modulation was ignored in the model.

@@

The parameters for the best model of the dynamic spectrum are given in Table \ref{tab:spec} and the resulting Gaussian model overlaid on Figure \ref{fig:spec}.  Two of the best-fit Gaussians are centered at the boundary of the 3~GHz band, so parameter estimates are actually limits. The other seven best-fit models appear contained by the 1~GHz wide band ($>90$\% of the modeled flux is within the 3~GHz band). In all cases, the typical burst spectrum has a FWHM of approximately 500~MHz. This confirms previous reports based on Arecibo and GBT data \citep{2016arXiv160308880S} with the wider (1~GHz) VLA bandwidth and extends this phenomenon to 3~GHz.

% ** signal or noise analysis offered by jim **

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57649.png}
 \end{minipage}
\caption{Spectra and best-fit model for nine bursts seen by the VLA from 2.5 to 3.5~GHz in the peak, dedispersed 5~ms integration. The solid line is a best-fit Gaussian model found through modeling. The typical flux density error per channel is 70~mJy, which is shown at the top right of each panel next to the MJD label.
\label{fig:spec}}
\end{center}
\end{figure*}

@@

\begin{figure*}[htb]
\begin{center}
\includegraphics[width=2\columnwidth]{corner57646}
\caption{A scatterplot matrix that shows the correlation of every pair of parameters in the MCMC run for burst 57646. The equivalent plot for the other bursts show normally-distributed samples, but this burst shows significant structure in the samples. The parameters ``amplitude'', ``center'', and ``$\sigma_{ch}$'' refer to the Gaussian spectral shape, while ``DM'', ``integration'', and ``W$_{int}$'' refer to the temporal shape. The parameters are given in units of channels (4 MHz) and integrations (5 ms).
\label{fig:corner}}
\end{center}
\end{figure*}

@@

Figure \ref{fig:ed} shows the \frb\ burst energy cumulative distribution as seen by the VLA and calculated from prior observations by Arecibo and the Green Bank Telescope \citep{2016Natur.531..202S,2016arXiv160308880S}. The latter two energy distributions are scaled from the fluence by assuming that all burst energy is included by the observation. This likely underestimates the burst energy for some bursts by a factor of a few, but should not affect the slope of the distribution.

@@

The burst-specific DM structure could be intrinsic to the emission mechanism or induced during propagation. Radio waves can be modified in a variety of ways (e.g., scintillation, scattering) and the duration of the emission in the source frame is not known \citep{2016arXiv160505890C}. \citet{2017arXiv170306580C} describe a model where plasma lensing near the source of an FRB can magnify radio emission by orders of magnitude. That model argues that the simultaneous appearance of a burst from 1.2 to 3.5~GHz requires a lens with the focal frequency be $\gtrsim 3.5$~GHz. This gives a constraint on the combination 
$$\frac{\rm DM_0}{a_{\rm AU}^2} \frac{d_{sl}d_{lo}/d_{so}}{1\, \rm kpc} \gtrsim (3.5 / 39.1)^2$$
where DM$_0$, $a$, and $d_{sl, lo, so}$\ are the electron column density of the lens, size of the lens, and lensing distances, respectively. A lens placed at $d_{sl}\approx1$~kpc is consistent with typical properties of the interstellar medium. Lenses that are influenced by the source would be much nearer to the source $d_{\sl} \ll 1$~kpc, requiring large values of $\rm{DM_0}/a_{\rm AU}^2$.

@@

Our refined analysis shows that \frb\ has even higher isotropic energies than previously reported for \frb, reaching energies $E_{\rm{max}}\approx10^{40}$\ erg. While this emission is coherent and beamed, the apparent energy is larger than Galactic analogues such as the giant pulses from the Crab pulsar \citep[$\sim10^{35}$~erg, $T_{b,\rm{Crab}}\sim10^{41}$~K;][]{2003Natur.422..141H,2014PhRvD..89j3009K}. Radio bursts from \frb\ require either dramatic scaling of known emission processes \citep{2016MNRAS.462..941L, 2016MNRAS.457..232C} or strong amplification by propagation effects \citep{2017arXiv170306580C}.

@@

@@

The chance of detecting a burst from \frb\ with the VLA changes substantially on day--month timescales. A major outstanding question is whether this time-variable rate is driven by an intrinsic process \citep{2016ApJ...826..226K} or extrinsic (propagation) effects in the host galaxy or intergalactic medium \citep{2017arXiv170306580C}. However, the temporal distribution of VLA detections alone shows that the \frb\ burst rate is significantly correlated on short timescales. That is, detecting a burst predicts that you are more likely to find another burst soon thereafter and not detecting a burst predicts you will likely not detect another burst soon thereafter.

@@

Using an assumed horizon redshift of 1, we estimate a rate of $R_{\rm{FRB}} \approx 5\times10^{-5} N_r^{-1} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$. This calculation ignores time dilation, which underestimates the burst rate by a factor of 2 at the redshift horizon. The narrow burst spectral structure seen for \frb\ also suggests that the FRB rate is understated by a factor of a few by ignoring bursts that fall outside the observing band. Finally, we note that this estimate assumes that Galactic latitude and propagation effects do not significantly affect the detectability of FRBs, although scintillation and scattering likely play significant roles \citep{2015MNRAS.451.3278M, 2017arXiv170306580C}.
